Probabilistic Policy Violation Detection with IPFWD

### 2017-02-17 paper/main.tex
Rough draft complete

  This approach is based on the premise that firewall performance can be
  improved by reducing the number of rules applied to each packet. Supposing a
  whitelist policy, a firewall has to apply every rule to each packet before
  it's denied.  With IPFWD, the system has a chance to accept any packet early
  and skip any further computation. Test results shows this reduces the system
  resources required to handle the same amount of traffic.

@@

    Firewalls restrict the types of incoming and outgoing network traffic
    according to a security policy. Without a security policy, administrators
    cannot know what to allow and disallow. Typical firewalls enforce the
    policy by blocking packets that do not match the rules. The default action
    for whitelist firewalls is to block all traffic, additional rules describe
    exceptions. Blacklist firewalls are the opposite in that by default they
    accept all traffic and additional rules describe packet types that are not
    allowed. Since predicting and blacklisting all possible attacks is
    infeasible, whitelist firewall rulesets are the accepted standard and the
    focus of this paper.

@@

  % How does it work?
  % - Probabilistic packet matching
  IPFWD's behavior is based on the assumption that the majority of network
  traffic is valid, malicious events are rare \cite{networktrafficanalysis},
  and it's sufficient to be notified of security policy violations instead of
  always stopping them \cite{TODO}.

@@

  This is especially valuable for a whitelist firewall rule set dealing with
  large amounts of rejected traffic; normally each rejected packet has all
  rules applied to it while IPFW looks for a match.  When no matches are
  encountered, the default rule is employed and the packet is rejected.

  In a typical data center environment the vast majority of network traffic is
  valid and going to be accepted eventually \cite{TODO}. However, these rule
  sets are often complex and take increasing time to match a given packet
  depending on the length. For a rule set of length $n$, we can expect to a
  packet to encounter $\frac{n}{2}$ failed matches before matching the correct
  rule and being accepted.

@@

        IPFWD works by reducing the average number of rules that fail to match
        before a packet is finished being processed. Very simple rulesets will
        gain little from reducing the number of rules applied to each packet.

@@

  % Talk about security policy enforcement
  Security policies define what is and is not allowed in an organization
  \cite{TODO}. At the network and host level, this is implemented in part by a
  firewall rule set. A policy violation on an internal network is more serious
  than an external network. Internet facing hosts can expect to be scanned,
  experience network anomalies and be attacked by third party agents
  semi-regularly.

@@

      Malicious network activity and network errors are likely to be repeated
      over time \cite{TODO}. The longer the activity persists, the more likely
      it is to be detected by IPFW even when IPFWD is allowing a high
      percentage of the network traffic through unchecked.

@@

    % Explain tests
    The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
    preserve CPU cycles for IPFW. Raw network throughput and CPU performance is
    the target metric, which can be difficult to assess with TCP
    \cite{TODO}.

@@

    % WWU 1G
    However, the WWU 1G tests provide an example of a NIC bound system. The
    lack of any performance gains in the WWU 1G tests provides an example of
    when IPFWD should not be used. In this system, the CPU and other system
    resources are more than sufficient to apply the entire IPFW ruleset to each
    packet at wire-speed. Reducing the amount of CPU time it takes to process
    each packet does not provide any benefit.

\section{Related Work}
  % Cisco and Juniper routers

  % Similiar work
  Related work \cite{exploitingpacketsampling, analysisnetflow,
  monitoringpacketsampling} into packet sampling in the space of network
  security has been focused on detecting and classifying threats, not
  performance enhancement.

  %Firewall decision trees? (Complete Redundancy Detection)

@@

  % Firewall rule reordering so we can "fail as early as possible"
  Ideally, a whitelist firewall's rule set would be constructed so that the
  most commonly applied matches are reached earlier rather than later. IPFWD
  accomplishes this task by artificial early stopping. Firewall rule reordering
  is a could be another solution to this problem. Given logs of common network
  traffic for the system, statistical analysis could show which rules are
  matched more often than others. With this information, the rule set could be
  reordered so that the most commonly applied rules are earliest.  TODO

  % creating baseline invalid packet % so we can extrapolate even when we don't
  % see any invalid packets

  % have detection of bad traffic reduce acceptance rate

@@

  In performance sensitive computing environments, complex firewall rules
  combined with high network throughput can result in heavy CPU load. IPFWD
  addresses this problem by providing a graceful tradeoff between packet
  visibility, total security policy enforcement and system performance.

  IPFWD reduces the average amount of work to process packets by adding an
  early rule to IPFD that has a chance to accept any packet. The security
  implications of this behavior have been addressed and the performance
  benefits provided have been demonstrated in a variety of systems.

  IPFWD can be directly applied to high network performance and resource
  conscious systems with complex firewall requirements to reduce CPU load and
  increase network throughput.

  This work also serves as a proof of concept for future probabilistic packet
  matching firewall implementations.

\appendices
\section{Additional Information}
  TODO link to github repository

>>>>

  This approach is based on the premise that firewall performance can be
  improved by reducing the number of rules applied to each packet. Supposing a
  white-list policy, a firewall has to apply every rule to each packet before
  it's denied.  With IPFWD, the system has a chance to accept any packet early
  and skip any further computation. Test results shows this reduces the system
  resources required to handle the same amount of traffic.

@@

    Firewalls restrict the types of incoming and outgoing network traffic
    according to a security policy. Without a security policy, administrators
    cannot know what to allow and disallow. Typical firewalls enforce the
    policy by blocking packets that do not match the rules. The default action
    for white-list firewalls is to block all traffic, additional rules describe
    exceptions. Blacklist firewalls are the opposite in that by default they
    accept all traffic and additional rules describe packet types that are not
    allowed. Since predicting and blacklisting all possible attacks is
    infeasible, white-list firewall rule sets are the accepted standard and the
    focus of this paper.

@@

  % How does it work?
  % - Probabilistic packet matching
  IPFWD's behavior is based on the assumption that the majority of network
  traffic is valid, malicious events are rare \cite{networktrafficanalysis},
  and it's sufficient to be notified of security policy violations instead of
  always stopping them. These assumptions will not hold in all environments,
  but we can leverage them for performance gains when they do.

@@

  This is especially valuable for a white-list firewall rule set dealing with
  large amounts of rejected traffic; normally each rejected packet has all
  rules applied to it while IPFW looks for a match.  When no matches are
  encountered, the default rule is employed and the packet is rejected.

  In a typical data center backend network the vast majority of traffic is
  valid and going to be accepted by the firewall's rule set. However, these
  rule sets are often complex and take increasing time to match packets
  depending on the number of rules. For a rule set of length $n$, we can expect
  to a packet to encounter $\frac{n}{2}$ failed matches before matching the
  correct rule and being accepted.

@@

        IPFWD works by reducing the average number of rules that fail to match
        before a packet is finished being processed. Very simple rule sets will
        gain little from reducing the number of rules applied to each packet.

@@

  % Talk about security policy enforcement
  Security policies define what is and is not allowed in an organization . At
  the network and host level, this is implemented in part by a firewall rule
  set. A policy violation on an internal network is more serious than an
  external network. Internet facing hosts can expect to be scanned, experience
  network anomalies and be attacked by third party agents semi-regularly.

@@

      Malicious network activity and network errors are likely to be repeated
      over time \cite{beacondetection}. The longer the activity persists, the
      more likely it is to be detected by IPFW even when IPFWD is allowing a
      high percentage of the network traffic through unchecked.

@@

    % Explain tests
    The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
    preserve CPU cycles for IPFW. Raw network throughput and CPU performance is
    the target metric, which can be difficult to assess with TCP
    \cite{netperf}.

@@

    % WWU 1G
    However, the WWU 1G tests provide an example of a NIC bound system. The
    lack of any performance gains in the WWU 1G tests provides an example of
    when IPFWD should not be used. In this system, the CPU and other system
    resources are more than sufficient to apply the entire IPFW rule set to
    each packet at wire-speed. Reducing the amount of CPU time it takes to
    process each packet does not provide any benefit.

\section{Related Work}

  % Packet sampling
  Other papers \cite{exploitingpacketsampling, analysisnetflow,
  monitoringpacketsampling} on packet sampling have been focused on detecting
  and classifying threats, not performance enhancement. However, their
  approaches to packet sampling and the implications addressed provided the
  idea for IPFWD.

  % IPFW performance
  The BSD community has produced several in depth investigations into network
  and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
  informed the tests run to measure IPFWD's performance benefits.

@@

  % Firewall rule reordering so we can "fail as early as possible"
  Ideally, a white-list firewall's rule set would be constructed so that the
  most commonly applied matches are reached earlier rather than later. IPFWD
  accomplishes this task by artificial early stopping. Firewall rule reordering
  is a could be another solution to this problem. Given logs of common network
  traffic for the system, statistical analysis could show which rules are
  matched more often than others. With this information, the rule set could be
  reordered so that the most commonly applied rules are earliest.

  Other work \cite{redundancydetection} has been done in removing redundant
  rules to increase firewall performance. It's possible that a combination of
  redundancy checking and priority reordering could provide further performance
  gains.

  % creating baseline invalid packet % so we can extrapolate even when we don't
  % see any invalid packets 
  Machine learning could be applied to packet samples to more accurately
  determine appropriate bounds on the early acceptance probability. Different
  systems and circumstances may require more or less stringent packet checking.
  Environments with very high confidence in network security might raise the
  lower bound of the acceptance rate to improve performance. Likewise, less
  confident environments may lower the upper bound to ensure an acceptable
  percentage of packets are always checked.

  % have detection of bad traffic reduce acceptance rate
  Future versions of IPFWD could benefit from additional modifiers to the early
  acceptance probability. For instance, the system could reduce the early
  acceptance probability whenever packets are dropped by the firewall under the
  assumption that invalid traffic likely precedes more invalid traffic.

@@

  In performance sensitive computing environments, complex firewall rules
  combined with high network throughput can result in heavy CPU load. IPFWD
  addresses this problem by providing a graceful trade off between packet
  visibility, total security policy enforcement and system performance.

  IPFWD reduces the average amount of work to process packets by adding an
  early rule to IPFW that has a chance to accept any packet. The security
  implications of this behavior have been addressed and the performance
  benefits provided have been demonstrated in a variety of systems.

  IPFWD can be directly applied to high network performance and resource
  conscious systems with complex firewall requirements to reduce CPU load and
  increase network throughput. This work also serves as a proof of concept for
  future probabilistic packet matching firewall implementations.

\appendices
\section{Additional Information}
  The source code for IPFWD, additional test results, and development cycle
  information may be found at \texttt{https://???}.
